\section{Substitution}

This section outlines how the Village decides how much to buy and
consume of goods that are partial substitutes.

\subsection{Defining substitutes}

A MaslowLevel may have more than one TradeGood, with relative weights,
like so:
\begin{verbatim}
level = {
  food = 0.19
  iron = 0.01
  mortality = 0.96
  max_work_fraction = 0.94
  name = "Barely fed"
}
\end{verbatim}

\subsection{Buying substitutes}

Let $x_i$ be the amount of good $i$, $y_i$ its price, and $w_i$ its
weight, normalised so their sum is 1. Then we require
\begin{eqnarray}
\label{eq:constraint}
\prod\limits_i x_i^{w_i} &=& C
\end{eqnarray}
where $C$ is the Village's consumption factor, and minimise the total
price paid
\begin{eqnarray*}
P = \sum\limits_i x_iy_i
\end{eqnarray*}
under this constraint. To find the amounts to bid for we use a
Lagrange multiplier:
\begin{eqnarray*}
\Lambda(\vec x, \vec y, \vec w) &=& \sum\limits_i x_iy_i -
\lambda\left(\prod\limits_i x_i^{w_i} - C\right)\\
\frac{\partial\Lambda}{\partial x_i} &=& y_i - \lambda w_ix^{w_i-1}\prod\limits_{j\ne i} x_j^{w_j}
\end{eqnarray*}
and setting the derivative to equal zero gives
\begin{eqnarray*}
\lambda w_ix_i^{-1}x_i^{w_i}\prod\limits_{j\ne i} x_j^{w_j} &=& y_i
\lambda &=& \frac{x_iy_i}{w_i}\prod\limits_i x_i^{-w_i}.
\end{eqnarray*}
We see that when equating the $\lambda$ for different values of $i$,
the product drops out, and (defining in passing an effective price
$y_i'=y_i/w_i$) we get
\begin{eqnarray}
\label{eq:equalprice}
x_iy_i' &=& x_jy_j'
\end{eqnarray}
for all $i, j$. Inserting this back into equation~\ref{eq:constraint},
we find
\begin{eqnarray}
x_1^{w_1}\prod\limits_{j\ne i} \left(\frac{x_1y_1'}{y_j'}\right)^{w_j} &=& C\\
x_1 &=& C \prod\limits_{j\ne i}\left(\frac{y_1'}{y_j'}\right)^{-w_j} \\
\label{eq:amount}
    &=& Cy_1'^{w_1-1} \prod\limits_{j\ne i}\left(y_j')^{w_j}
\end{eqnarray}
which looks fearsome but is straightforward to calculate; and from
there we can get the other $x_i$ by inserting into
equation~\ref{eq:equalprice}.

A sanity check: Suppose two goods have equal weights 0.5, and equal
prices 1; and suppose $C$ is 1000. Then $x_1$ is just 1000 and so is
$x_2$, which is sensible - equal goods, equal amounts, and equal to
the consumption capacity. Now suppose the first good has weight
two-thirds, and the prices are still equal. Then $x_1$ becomes 1260
and $x_2$ is 630, so the higher-weighted good is used more and the
total is still very close to 2000. Conversely if $y_1=0.75$ and
$y_2=1.5$, giving the same effective prices, then $x_1$ becomes 1414
and $x_2$ 707. These all seem like sensible numbers.

% (* (* 1000 (pow 1.5 -0.5) (pow 3 0.5)) 1.5 (/ 1.0 3))

\subsection{Consuming substitutes}

In this case there is no price to guide us. The constraint equation is
the same, since it reflects that the total consumption should be some
given constant. But the market prices don't necessarily reflect the
real consumption costs. So the new problem is to satisfy equation~\ref{eq:constraint}
while minimising the total fractional consumption $\sum\limits_i
x_i/X_i$ where $X_i$ is the reserve of good $i$. We see that this is
equivalent to just setting prices equal to inverse reserves, and we
can reuse equations~\ref{eq:equalprice} and \ref{eq:amount} with the
substitution $y_i' = \frac{1}{w_iX_i}$. By inspection we see that
consumption is directly proportional to weight and reserves, which
makes sense. To sanity-check the result, suppose we have equal
amounts, say 1000, of each good, equal weights, and consumption
capacity also 1000. Then $x_1$ is 1000 and so is $x_2$. Changing the
weights, as before, to two-thirds and one-third, we get 1260 and 630,
as before. Which is really bad because we assumed we had 1000 of
each. So we will have to insert an additional clause: Don't use more
than you have; if forced to attempt it, use everything available and redo
the problem with one less good. However, on the assumption that we
have 2000 of each, we get $x_1$ of 1260 again and $x_2$ of 630, so
it's fine. If we have only 1000 of good 2, then we get 1587 and 397,
which is also ok.

%(let ((X1 2000)
%      (c1 1000)
%      (X2 1000) 
%      (w1 0.6666) 
%      (w2 (- 1 w1))
%      (x1 (* c1 (pow (/ 1.0 (* X1 w1)) (- w1 1)) (pow (/ 1.0 (* X2 w2)) w2)))
%      (x2 (* w2 X2 (/ 1.0 w1 X1))))
%(* x1 x2))

